---
title: "Distribution of wild animals "
output: word_document
date: "2024-03-05"
---
# התקנת חבילה קריאה מקובץ אקסל
install.packages("readxl")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries
```{r libraries, include=TRUE,echo=TRUE}
library(fitdistrplus)
library(magrittr)
library(readxl)
```


```{r generate}
dataset <- read_excel(file.choose())

# Convert the "datasets חיות בר" column to POSIXct format
dataset$`חיות בר` <- as.POSIXct(dataset$`חיות בר`, format = "%Y-%m-%d %H:%M:%S")

# Calculate the time difference between consecutive rows in the "datasets חיות בר" column in minutes
time_diff <- diff(dataset$`חיות בר`, units = "mins")

# Add a new column named "Time DiffMinutes" to the dataset and populate it with the time difference in minutes
dataset$Time.DiffMinutes <- c(0, time_diff / 60)

# Print the "Time DiffMinutes" column
print(dataset$Time.DiffMinutes)
```

## First Impressions 

For that we can use the following plots:
```{r empirical plots}
par(mfrow=c(1,3)) # defines 2 plots on the same window
plot(dataset$Time.DiffMinutes,ylab="Sample values") # plot ( left graph )
hist(dataset$Time.DiffMinutes, ylab="Frequency ") # histogram (right graph)
boxplot(dataset$Time.DiffMinutes)
```


## According to the plots shown, we chose to check a normal distribution

```{r fitting norm}
normFit<-fitdist(dataset$Time.DiffMinutes,"norm") # fitting a normal distrbution 
summary(normFit) # summary function displays the results 

```
The estimate are R estimates of the paramters of a normal distribution - mean and sd.

```{r results="hide"}
normFit$estimate[1]  # is the first paramter- mean
normFit$estimate[2]  # is the second paramter- sd
```

Now we will use plotdist to plot our data versus the theoretical distribution 

```{r}
plotdist(dataset$Time.DiffMinutes,"norm",para=list(mean=normFit$estimate[1],sd=normFit$estimate[2])) 

```

## Now we have chosen to test an exponential distribution

```{r fitting norm}
expFit<-fitdist(dataset$Time.DiffMinutes,"exp") # fitting a exponential distrbution 
summary(expFit) # summary function displays the results 


```

We can reach the paramter directly from code using the following:

```{r results="hide"}
expFit$estimate[1]  #rate
```

Now we will use plotdist to plot our data versus the theoretical distribution 

```{r}
plotdist(dataset$Time.DiffMinutes,"exp",para=list(rate=expFit$estimate[1])) 


```

